---
title: "Visualising Trophic Networks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualising Trophic Networks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
## Figure options
knitr::opts_chunk$set(
  fig.width = 8,
  fig.height = 6,
  fig.align = "center",
  dpi = 150
)

## Load libraries
library(foodwebbuilder)
```

# Overview

This vignette shows how to:  
1. Build a size-structured trophic metaweb from the example datasets shipped with `foodwebbuilder`.  
2. Compute a few simple network metrics (trophic level, trophic breadth, degrees, bottom-up fluxes).  
3. Visualise networks using Cartesian and radial layouts.  

**Convention used throughout:**  
the metaweb adjacency matrix is organised as *rows = prey/resources and columns = consumers/predators*.

# Load Data

```{r}
data(ind_measure)
data(fish_diet_shift)
data(resource_diet_shift)
data(pred_win)

head(ind_measure)
head(fish_diet_shift)
head(resource_diet_shift)
head(pred_win)
```

# Build the Metaweb

```{r}
NUM_CLASSES <- 3
SELECTED_RESOURCES <- c("det", "biof", "phytob", "macroph",
                        "phytopl", "zoopl", "zoob")

# Keep only species that exist consistently across inputs
ind_clean <- remove_missing_species(ind_measure,
                                      fish_diet_shift,
                                      pred_win)

# Build size classes (species x size classes)
tab_size_classes <- compute_size_classes(ind_clean, NUM_CLASSES)

# Build metaweb (rows = prey/resources, cols = consumers/predators)
metaweb <- build_metaweb(tab_size_classes,
                         pred_win,
                         fish_diet_shift,
                         resource_diet_shift,
                         NUM_CLASSES,
                         SELECTED_RESOURCES,
                         method_resource_fish    = "midpoint",
                         method_predation_window = "midpoint",
                         method_fish_fish        = "midpoint"
                         )
```

# Visualise the metaweb matrix

```{r}
mat <- as.matrix(metaweb)

old_par <- par(no.readonly = TRUE)
on.exit(par(old_par), add = TRUE)

par(mar = c(2, 10, 10, 2), xpd = NA)

z <- mat[nrow(mat):1, , drop = FALSE]
pal <- grDevices::colorRampPalette(c("blue", "white", "red"))(100)

graphics::image(
  x = seq_len(ncol(mat)),
  y = seq_len(nrow(mat)),
  z = t(z),
  col = pal,
  axes = FALSE,
  xlab = "",
  ylab = ""
)

graphics::box()

# ticks only
graphics::axis(3, at = seq_len(ncol(mat)), labels = FALSE, tck = -0.015)
graphics::axis(2, at = seq_len(nrow(mat)), labels = FALSE, tck = -0.015)

# draw all labels manually
graphics::text(
  x = seq_len(ncol(mat)),
  y = nrow(mat) + 0.8,
  labels = colnames(mat),
  srt = 90, adj = 0, cex = 0.55
)

graphics::text(
  x = -0.3,
  y = seq_len(nrow(mat)),
  labels = rev(rownames(mat)),
  adj = 1, cex = 0.65
)

graphics::mtext("Global Metaweb", side = 3, line = 7.5, font = 2, cex = 1.2)
graphics::mtext("Consumers / Predators (columns)", side = 3, line = 3.5, font = 2)
graphics::mtext("Prey / Resources (rows)", side = 2, line = 4.5, font = 2)
```

# Compute network metrics

```{r}
# Structural metrics
basal_nodes <- get_basal_nodes(metaweb)
leaf_nodes  <- get_leaf_nodes(metaweb)

degree_in  <- compute_inward_degree(metaweb)
degree_out <- compute_outward_degree(metaweb)

TL <- compute_trophic_level(metaweb)
TB <- compute_trophic_breadth(metaweb, TL)

# Fluxes (matrix of accumulated bottom-up fluxes)
fluxes <- compute_bottom_up_fluxes(metaweb)
fluxes

incoming_flux <- colSums(fluxes)  # received by consumers (columns)
outgoing_flux <- rowSums(fluxes)  # emitted from prey/resources (rows)

op <- par(no.readonly = TRUE)
on.exit(par(op), add = TRUE)

par(mfrow = c(1, 2), mar = c(5, 10, 2, 1))

barplot(sort(incoming_flux, decreasing = TRUE),
        horiz = TRUE, las = 1, cex.names = 0.6,
        main = "Incoming flux per node")

barplot(sort(outgoing_flux, decreasing = TRUE),
        horiz = TRUE, las = 1, cex.names = 0.6,
        main = "Outgoing flux per node")

#barplot(fluxes, xlab = "Biomass Flux", horiz = TRUE, las = 1)

```

# Network visualisations
## Degree-based & Trophic metrics layout

```{r}
plot_network(
  metaweb,
  x = degree_out,
  y = degree_in,
  labels = colnames(metaweb),
  xlab = "Outward degree (row sums)",
  ylab = "Inward degree (column sums)",
  label_space_x = 0.10
)

plot_network(
  metaweb,
  x = TB,
  y = TL,
  labels = colnames(metaweb),
  xlab = "Trophic breadth (sd of resource TL)",
  ylab = "Trophic level",
  add_legend = "bottomright",
  label_space_x = 0.10
)
```

## Flux visualisation
When plotting 'fluxes', edge widths reflect flux magnitude (relative to the maximum absolute value).

```{r}
plot_network(
  fluxes,
  x = degree_out,
  y = degree_in,
  labels = colnames(metaweb),
  xlab = "Outward degree",
  ylab = "Inward degree",
  label_space_x = 0.10,
  line_width_max = 2
)

plot_network(
  fluxes,
  x = TB,
  y = TL,
  labels = colnames(metaweb),
  xlab = "Trophic breadth",
  ylab = "Trophic level",
  add_legend = "bottomright",
  label_space_x = 0.10,
  line_width_max = 2
)
```

## Radial Layout

```{r}
plot_network_radial(
  fluxes,
  x = TL,
  labels = colnames(metaweb),
  rotations = 0.975,
  line_width_max = 1
)

plot_network_radial(
  fluxes,
  x = degree_in,
  labels = colnames(metaweb),
  rotations = 0.5,
  scale = 1.25,
  line_width_max = 1
)

plot_network_radial(
  fluxes,
  x = TL,
  y = TL,
  labels = colnames(metaweb),
  rotations = 1.5,
  scale = 1.25,
  line_width_max = 1
)
```

```
```

